Discovery of Rotational Modulations in the Planetary-mass Companion 2M1207\lowercase{b}: An intermediate Rotation Period and Heterogeneous Clouds in a Low Gravity Atmosphere

### 2015-09-25 summary/letter/letter_vAF.tex
submitted

\title{Discovery of Rotational Modulations in the Planetary-mass
  Companion 2M1207b: A Slow Rotation Period and Heterogeneous Clouds in a Low
  Gravity Atmosphere}
\shorttitle{Variability of 2M1207 b}
\author{Yifan Zhou\altaffilmark{1}, D\'aniel Apai\altaffilmark{1,2,3},
  Glenn Schneider\altaffilmark{1},  Mark S. Marley\altaffilmark{4}}

\altaffiltext{1}{Department of Astronomy/Steward Observatory, The
  University of Arizona, 933 N. Cherry Ave., Tucson, AZ, 85721, USA,
  \href{mailto:yifzhou@email.arizona.edu}{yifzhou@email.arizona.edu}
}
\altaffiltext{2}{Lunar and Planetary Laboratory, The University of
  Arizona, 1640 E. University Blvd., Tucson, AZ 85718, USA}
\altaffiltext{3}{Earths in Other Solar Systems Team, NASA Nexus for
  Exoplanet System Science}
\altaffiltext{4}{NASA Ames Research Center, Naval Air Station,
  Moffett Field,Mountain View, CA 94035, USA}

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, periodic light curve directly probes the rotational periods of ultra-cool objects.

  We present here, for the first time, time-resolved high-precision
  photometric measurements of a planetary-mass companion, 2MASS1207b,
  to a brown dwarf primary.  We observed the binary system with
  HST/WFC3 in two bands and with two spacecraft roll angles. Using
  point spread function-based photometry we reach a nearly
  photon-noise limited photometric accuracy for both components. While
  the primary is consistent with a flat lightcurve, the secondary
  shows modulations that are clearly detected in the combined
  lightcurve as well as in different subsets of the data.  The
  amplitudes are 1.45\% in the F125W and 0.92\% in the F160W filters;
  we find a consistent period of $10.2^{+0.9}_{-0.8}$ h and similar
  phase in both bands. The J- and H-band amplitude ratio of 2M1207 bis very
  similar to a field brown dwarf that has identical spectral type but
  different J-H color.  Importantly, our study also measures, for
  the first time, the rotational period for directly an imaged
  planetary-mass companion.
\end{abstract}

\keywords{brown dwarfs -- planets and satellites: atmospheres -- planets
  and satellites: individual (2M1207 b) -- techniques: photometric}
\maketitle
%
\section{Introduction}

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}
  \caption{Point spread function subtraction allows isolating the
    secondary and accurately measuring its brightness in our WFC3
    F160W images. {\em Upper}: original image, the position of 2M1207
    b is indicated by a white circle.  {\em Lower:} residual image --
    after the subtraction of the hybrid PSF, 2M1207 b is detected at a
    high significant level.}
  \label{fig:1}
\end{figure*}

Presence of condensate clouds is one the most unique features of the
ultra-cool atmosphere of directly imaged exoplanet and brown
dwarfs. Studies of formation and properties of condensate clouds
\citep[e.g.][]{Ackerman2001, Burrows2006a, Helling2008, Allard2012}
have achieved great improvement in understanding the cloud behavior
across different spectral types, especially in the explanation of L-T
transition \citep[e.g.][]{Burrows2006a, Marley2010}.  Surface gravity
is suggested to be the second key parameters in defining cloud
structures \citep[e.g.][]{Marley2012} after effective temperature. Low
surface gravity objects (e.g. HR8799 bcd, \cite{Marois2008a}, 2M1207
b, \cite{Chauvin2004}) are significantly redder and under-luminous
compared to brown dwarfs.  The anomalous color and luminosity of low surface gravity
objects supports model including unusually thick clouds
\citep{Currie2011, Madhusudhan2011,Skemer2011, Skemer2012}. However, due to lack of observational
constraint, the dependence of cloud properties on surface gravity is
not very well modeled.

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
\citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2014,Metchev2015,Heinze2015}. \cite{Kostov2013}
demonstrated that these techniques can be applied to directly imaged
exoplanets, too, allowing comparative studies of objects with
different surface gravities. However, high contrast magnifies the
challenges for directly imaged exoplanets and planetary-mass
companions to acquire high-precision light curves comparing to brown
dwarfs.

  2M1207 b \cite{Chauvin2004} is the first directly imaged
planetary-mass companion. \cite{Chauvin2005,Song2006} confirmed that 2M1207 b and
its host 2M1207 A form a bound, co-moving system. 2M1207 A and
b have an angular separation of $0.78''$, which is corresponding to a
projected separation of 41.2 AU at a distance of 52.4 pc
\citep[e.g.][]{Ducourant2008}. Combining 2M1207 b's age and near infrared
luminosity with  brown dwarf cooling models
\citep[e.g.][]{Baraffe2003}, the object's mass is estimated to be
2.3-4.8 M$_{\mathrm{Jup}}$ \citep{Barman2011b}. Even though a
circumsubstellar disk was discovered around 2M1207A
\citep[][]{Sterzik2004}, the high companion-to-host
mass ratio and large separation argue for binary-like gravitational fragmentation
formation\citep{Mohanty2007}.

Early observations revealed that 2M1207 b's color is much redder and
its near-infrared luminosity is much lower than those of field brown
dwarfs with similar spectra\citep[e.g][]{Mohanty2007, Skemer2011,
  Barman2011b}. 2M1207b's luminosity -- as derived from near-infrared
photometry -- is $\sim2.5$ mag lower than that predicted based on its
mid- to late L spectral type and effective temperature of $\sim 1600$
\citep{Patience2010}.  Based multi-band near-infrared photometry
\citet{Skemer2011} argued that the apparent under-luminosity of 2M1207
b may be explained by a model of a spatially heterogeneous atmosphere
composed of patches of thin and patches of unusually thick clouds.
Similarly, \cite{Barman2011b} argued that non-local chemical
equilibrium could play an equally important role as thick clouds in
defining 2M1207b's color and luminosity.

@@

In this {\em Letter} we present the first, high-cadence,
high-precision, time-resolved {\em Hubble Space Telescope} (HST)
photometric time series of 2M1207 b, a directly imaged planetary-mass
object. We successfully detect rotational modulation and measure the
amplitudes in two bands and determine the rotational period. These observations probe the spatial heterogeneity and
vertical structure of clouds in planetary mass objects for the first
time.

@@

We obtained direct images of the 2M1207A+b system on UT 2014 April 11
from 08:07:47 to 16:53:18 using HST and its Wide Field Camera 3
\citep[WFC3, pixel scale=$0.130$mas/pixel, ][]{Mackenty2008} in the frame of
the HST Program GO-13418 (PI: D. Apai). We acquired the observations
in filters F125W ($\lambda_{\mathrm{pivot}}$ = 1245.9 nm, full width
at half maximum (FWHM) = 301.5 nm) and F160W
($\lambda_{\mathrm{pivot}}$ 1540.52, FWHM = 287.9 nm), roughly
corresponding to the J and H bands. We used the $256\times256$ pixels
sub-array mode to avoid memory dumps during the observations.  In
order to provide a near-continuous coverage for detecting modulations
we observed the 2M1207 system in 6 consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58 minutes long Earth
occultations every 94 minutes.

The observations applied space craft rolls each two orbit to allow
roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbit 1, 3, and 5, and those for 2, 4, and 6
differ by $25^{\circ}$. At the separation of 2M1207 b, this angle
difference corresponds to a displacement of $0.34''$, or 2.75 and 2.30
resolution elements in F125W and F160W, respectively. In each orbit we
took 8 SPARS10 exposure sequence with NSAMP=10, alternating between
F160W and F125W filters, with 2--3 identical exposures in each
exposure sequence. To improve sampling and reduce the risk that the
core of point spread function (PSF) is affected by bad pixels, we
applied a 4-point dither pattern with differential "X/Y" offsets of
1.375" in the detector frame, providing optimal non-integral (half
pixel) step of 10.5 and 8.5 pixels in F125W and F160W,
respectively. In total, we obtained 70 images with 10 non-destructive
read-outs in F125W and 64 images in F160W with exposure time of 88.4~s
for both filters.

@@

We started the reduction from the \flt{} produced by the WFC3's
\texttt{calwfc3} pipeline. We did not opt to use \ima{} that contain
all non-destructive read-outs, because they provided less information
on 2M1207A, which saturated after the first few samples.  The \flt{}
are results of basic calibration, including dark current correction,
non-linearity correction, flat field correction, as well as
up-the-ramp fit on the non-destructive read-outs. 
Pixels with data quality flags ``bad detector pixels'', ``unstable
response'', and ``bad or uncertain flat value'' were masked out and
excluded from further analysis as suggested by previous transit
exoplanet spectroscopic observations\citep[e.g.][]{Berta2012,
  Kreidberg2014}.

The major challenge of high contrast observation with WFC3/IR is the
fact that the detector is significantly under-sampled.  2M1207 A and b
are only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF on the
detector. When
applying roll subtraction, notable artifacts are introduced by image
shifting and interpolation.  \tinytim{} PSF
simulator\citep{Krist1995} offers a solution by providing Nyquist or
better sampled PSF, but systematic errors of \tinytim{} PSF for WFC3
limits its ability in high precision photometry\citep{Biretta2014}.
Building on the large number of PSFs obtained in our program at two
different roll angles, we followed a novel, two-step approach that
uses a hybrid PSF, {\color{blue}which combined the strengths of the roll
subtraction-based and the simulated PSF-based approaches.}
First, based on our observations we observationally derived correction maps
for \tinytim{} that accurately described the scattered light component
for the primary at the correct location on the detector. Second, we
carried out a PSF-photometry using hybrid PSFs composed by \tinytim{}
PSFs and the correction map  by simultaneously
minimizing the residuals from the primary and the secondary.
% However, we are able to fully characterize the difference of model
% and observed PSFs with 6 orbits time-resolved observation data. To
% obtain robust \tinytim{} PSF photometry, we design a 2-round PSF
% fitting strategy: 1. calculating correction map for \tinytim{};
% 2. hybrid PSF photometry.

For both of 2 steps, we used \tinytim{} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra (2M1207A:
\cite{Bonnefoy2014}, 2M1207 b: \cite{Patience2010}), the telescope's
actual focus, and the telescope jitter.  We used the set of \tinytim{}
parameters provided by \cite{Biretta2014} to improve modeling the cold
mask, diffraction spikes, and the coma. The focus parameters are are
interpolated to the precise time of the observations using the
tabulated values provided by
STScI\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}.
To align the \tinytim{} PSF to the observed PSF of 2M1207 A, we
moved the over-sampled PSF on a coordinate grid (gird size=0.001 pixel)
using cubic interpolation, and searched for the position that minimizes
the rms difference of the observed and the re-binned \tinytim{} PSF over
a region centered on 2M1207 A with a 5-pixel-radius aperture centered
on 2M1207 b excluded.  Then we introduced another \tinytim{} PSF for
2M1207 b and fit the position of 2M1207 b and the scales of the
\tinytim{} PSFs of 2M1207 A and b simultaneously by minimizing the
residual from both primary and secondary. In the first step, we discovered that the
difference of observed PSFs and model PSFs were very stable for a
specified telescope roll angle and dithering position. Therefore, at the end of the first step,
we derive 8 (2 roll angles $\times$ 4 dithering positions) correction
maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ was a combination of two scaled \tinytim{} PSFs
for 2M1207 A and b. In the second round, we combined the correction
map linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and scaled the correction map together with the two PSFs so that the
residual which is expressed as
\begin{equation}
  \mathrm{residual} = \mathrm{Image} - a\cdot \mathrm{PSF_{A}} - b\cdot
    \mathrm{PSF_{b}}- c\cdot\mathrm{ Corr}
  \end{equation}
  is minimized by least square fitting.
We found that by introducing the correction term,
the reduced $\chi^{2}$ was decreased from $\sim 10$ to
$\sim 1$. Relative photometry was acquired from the scaling
parameters of the \tinytim{} PSFs.

Our final step was to correct for a slight apparent trend between the
position of the targets on the detector and their flux. We attributed
this to a combination of slight changes in the PSF profile due to
pixelation and to the effect of imperfectly corrected pixel-to-pixel
sensitivity variations. We corrected for these apparent
position-dependent flux changes by normalizing each photometric point
by the median of all fluxes measured when the target was at the
same position, i.e. combining data over 6 orbits.  We note that this
correction is small and, as we demonstrate in the next sections,
cannot introduce artificial modulations that resemble the long-period
variations we identify in 2M1207b.

% PSF profiles change with exposure positions due to
% pixelation, especially for the case that WFC3 IR is significantly
% under-sampled. Also, the flat fields may potentially have large scale
% structures \citep{dressel2012wide}. We find a correlation of
% photometry with PSF positions on detector frame for both 2M1207 A and
% b. Correction is made by normalizing each group of
% exposures that have the same dithering position and telescope roll angle
% individually -- we take the median of the fluxes that are measured
% from these exposures as normalization factors and divided them from
% every photometric measurement. Because the normalization factor for
% each group of exposures is calculated across the whole observation in
% time domain,
% this step has negligible impacts on variability
% analysis.

@@

First we estimated the photon noise for the photometry of
2M1207 b. The total photon noise of the photometry was calculated by
combining the photon noise of every pixel, which was derived
from count rates and detector gain. The photon noises 
in F125W and F160W are 1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207 A and b were fitted simultaneously, the
uncertainties for photometry and position of the primary and secondary
were coupled. Errors in position measurements of 2M1207 A could
potentially affect the photometry of 2M1207 b. We used a Monte Carlo
(MC) method to evaluate the overall systematic of the PSF fitting. We
applied photometry to images that were added with random
Poisson noise and repeated the photometry procedure for 1000 times. The
uncertainties for F125W and F160W photometry were found to be 1.34\% and 1.12\%,
respectively.

@@

A further contribution to photometric uncertainties may be introduced
by imperfectly corrected pixel-to-pixel sensitivity
differences. 2M1207 b were observed at 8 different positions on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction could introduce position-dependent
differences in the count rates. The uncertainty of WFC3 IR flat field
is typically $\sim 1\%$ \citep{dressel2012wide}.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
    variability verification tests. Individual measurements are
    plotted with gray crosses. Photometric measurements of the same exposure
    sequence are binned, and binned photometry are plotted with points
    or squares. Best fitted sinusoidal curves are plotted with solid
    lines. {\em Upper}: binned measurements taken in dithering
    position 1 and 3 (red points) and that taken in 2 and 4 (blue
    squares) are plotted with different symbols. They demonstrate same
    trend of modulation. In upper left panel, green line is sinusoidal
    wave fitted with all parameter set free, and purple line is
    sinusoidal wave fitted with period set the same as that of
    F125W. {\em Middle}: sinusoidal waves fitted without using the
    data taken in Orbit \#1. These curves are almost identical to the
    curves plotted in upper panel. {\em Lower}: photometry measured
    with AFEM-added images and best fitted sinul curves. These
    points and curves are also almost identical to those plotted
    in the upper panel.}
  \label{fig:2}
\end{figure*}

@@

The light curves that resulted from our photometry showed apparently
sinusoidal modulations, discussed in more details in
\S,\ref{Results}. To verify that these modulations are intrinsic to
the object and not results of our data reduction procedures or due to
instrumental changes, we carried out three different tests.

@@

As a second test, we repeated the analysis neglecting the first
orbit. The motivation behind this test is that, due to spacecraft
thermal settling, the first orbits of HST observations are often
slightly unstable, and neglected in high-precision studies
\citep[e.g.][]{Mandell2013}. Indeed, in our analysis 2M1207 A is
significantly fainter in the first orbit (Figure \ref{fig:3}) than in
the subsequent ones.  Our analysis based on orbits 2--6 found
essentially identical results to our analysis using the whole 6 orbits, based on
which we conclude that the first less reliable orbit does not affect
our results significantly (Figure \ref{fig:2}, middle panel).

As a third test, we explored whether a subset of images, perhaps due to
imperfectly normalized or correlated with specific instrument states,
could drive the light curves into an apparently sinusoidal shape. To
test this possibility, we split the data into two temporally
overlapping halves: subset one were images taken at dithering position
1 and 3, and subset two were those taken at dithering position 2 and
4. For both subsets, we repeated our analysis independently.  For both
of F125W and F160W, two halves demonstrated similar sinusoidal
modulations.  Our analysis detect sinusoidal modulations in {\em both}
subsets and in {\em both} filters, with periods and amplitudes
consistent with those derived from the complete data set (Figure
\ref{fig:2}, upper panel).

@@

We used a MC method to analyze the light curve, and provided amplitudes
and periods as well as their uncertainties for both filters. We
generated series of Gaussian distributed random noise with standard deviation
same as the photon noise, add the noise to the original light curves, and
fit sinusoids to the newly combined light curves. We repeated these
procedures  for 100,000 times and obtain the distributions of period
and amplitude (Figure \ref{fig:4}).

>>>>

\title{Discovery of Rotational Modulations in the Planetary-mass
  Companion 2M1207\lowercase{b}: An intermediate Rotation Period and Heterogeneous Clouds in a Low
  Gravity Atmosphere}
\shorttitle{Variability of 2M1207b}
\author{Yifan Zhou\altaffilmark{1}, D\'aniel Apai\altaffilmark{1,2,3},
  Glenn H Schneider\altaffilmark{1},  Mark S. Marley\altaffilmark{4},
  Adam P. Showman\altaffilmark{2}}

\altaffiltext{1}{Department of Astronomy/Steward Observatory, The
  University of Arizona, 933 N. Cherry Ave., Tucson, AZ, 85721, USA,
  \href{mailto:yifzhou@email.arizona.edu}{yifzhou@email.arizona.edu}
}
\altaffiltext{2}{Department of Planetary Science/Lunar and Planetary Laboratory, The University of
  Arizona, 1640 E. University Blvd., Tucson, AZ 85718, USA}
\altaffiltext{3}{Earths in Other Solar Systems Team, NASA Nexus for
  Exoplanet System Science}
\altaffiltext{4}{NASA Ames Research Center, Naval Air Station,
  Moffett Field,Mountain View, CA 94035, USA}

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, periodic light curves directly probe the
  rotational periods of ultra-cool objects.  We present here, for the
  first time, time-resolved high-precision photometric measurements of
  a planetary-mass companion, 2M1207b.  We observed the binary
  system with HST/WFC3 in two bands and with two spacecraft roll
  angles. Using point spread function-based photometry, we reach a
  nearly photon-noise limited accuracy for both the primary and the
  secondary. While the primary is consistent with a flat light curve,
  the secondary shows modulations that are clearly detected in the
  combined light curve as well as in different subsets of the data.
  The amplitudes are 1.45\% in the F125W and 0.92\% in the F160W
  filters, respectively. We find a consistent period of
  $10.2^{+0.9}_{-0.8}$ hours and
  similar phases in both bands. The J- and H-band amplitude ratio of
  2M1207b is very similar to a field brown dwarf that has identical
  spectral type but different J-H color.  Importantly, our study also
  measures, for the first time, the rotation period for a directly 
  imaged planetary-mass companion to another star.
\end{abstract}

\keywords{brown dwarfs -- planets and satellites: atmospheres -- planets
  and satellites: individual (2M1207b) -- techniques: photometric}
\maketitle
%
\section{Introduction}

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}
  \caption{Point spread function subtraction allows isolating the
    secondary and accurately measuring its brightness in our WFC3
    F160W images. {\em Upper}: original image, the position of 2M1207
    b is indicated by a white circle.  {\em Lower:} residual image --
    after the subtraction of the hybrid PSF, 2M1207b is detected at a
    high significance level.}
  \label{fig:1}
\end{figure*}

Presence of condensate clouds is one the most unique features of the
ultra-cool atmosphere of directly imaged exoplanet and brown
dwarfs. Studies of formation and properties of condensate clouds
\citep[e.g.][]{Ackerman2001, Burrows2006a, Helling2008, Allard2012}
have made great progress on the cloud behaviors across different
spectral types, especially the role that clouds play in the L-T
transition \citep[e.g.][]{Burrows2006a, Marley2010}.  Surface gravity
is suggested to be the second key parameters in defining cloud
structures \citep[e.g.][]{Marley2012} after effective temperature. Low
surface gravity objects (e.g. HR8799 bcd, \cite{Marois2008a}, 2M1207b, \cite{Chauvin2004}) are significantly redder and under-luminous
compared to field brown dwarfs.  The anomalous color and luminosity of
low surface gravity objects support models including unusually thick
clouds \citep{Currie2011, Madhusudhan2011,Skemer2011,
  Skemer2012}. However, due to lack of observational constraint, the
dependence of cloud properties on surface gravity is not very well
modeled.

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping \citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2015,Metchev2015,Heinze2015}. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
. \citet{Kostov2013}
demonstrated that these techniques can be applied to directly imaged
exoplanets, too, allowing comparative studies of objects with
different surface gravities. However, high contrast magnifies the
challenges for directly imaged exoplanets and planetary-mass
companions to acquire high-precision light curves comparing to brown
dwarfs.

2M1207b \citet{Chauvin2004} is the first directly imaged extra-solar
planetary-mass companion. \citet{Chauvin2005} and \citet{Song2006}
confirmed that 2M1207b and its host 2M1207A form a bound, co-moving
system. 2M1207A and b have an angular separation of $0.78''$, which
corresponds to a projected separation of 41.2 AU at a distance of 52.4
pc \citep[e.g.][]{Ducourant2008}. Combining 2M1207b's age and near
infrared luminosity with brown dwarf cooling models
\citep[e.g.][]{Baraffe2003}, the object's mass is estimated to be
2.3-4.8 M$_{\mathrm{Jup}}$ \citep{Barman2011b}. Even though a
circumsubstellar disk was discovered around 2M1207A
\citep[][]{Sterzik2004}, the high companion-to-host mass ratio and
large separation argue for binary-like gravitational fragmentation
formation\citep{Mohanty2007}.

Early observations revealed that 2M1207b's color is much redder and
its near-infrared luminosity is much lower than those of field brown
dwarfs with similar spectra\citep[e.g][]{Mohanty2007, Skemer2011,
  Barman2011b}. 2M1207b's luminosity -- as derived from near-infrared
photometry -- is $\sim2.5$ mag lower than that predicted based on its
mid- to late L spectral type and effective temperature of $\sim 1600$
 K\citep{Patience2010}.  Based on multi-band, near-infrared photometry,
\citet{Skemer2011} argued that the apparent under-luminosity of 2M1207
b could be explained by a model of a spatially heterogeneous atmosphere
composed of patches of thin and patches of unusually thick clouds.
Similarly, \cite{Barman2011b} argued that non-local chemical
equilibrium could play an equally important role as thick clouds in
defining 2M1207b's color and luminosity.

@@

In this {\em Letter} we present the first, high-contrast, high-cadence,
high-precision, time-resolved {\em Hubble Space Telescope} (HST)
photometric time series of 2M1207b, a directly imaged planetary-mass
object. We successfully detect rotational modulation and measure the
amplitudes in two bands and determine the rotational period. These
observations probe the spatial heterogeneity and vertical structure of
clouds in planetary mass objects for the first time.

@@

We obtained direct images of the 2M1207A+b system on UT 2014 April 11
from 08:07:47 to 16:53:18 using HST and its Wide Field Camera 3
\citep[WFC3, pixel scale=$0.130$mas/pixel, ][]{Mackenty2008} in the frame of
the HST Program GO-13418 (PI: D. Apai). We acquired the observations
in filters F125W ($\lambda_{\mathrm{pivot}}$ = 1245.9 nm, full width
at half maximum (FWHM) = 301.5 nm) and F160W
($\lambda_{\mathrm{pivot}}$ 1540.52, FWHM = 287.9 nm), roughly
corresponding to the J and H bands. We used the $256\times256$ pixels
sub-array mode to avoid memory dumps during the observations.  In
order to provide a near-continuous coverage for detecting modulations
we observed the 2M1207 system in 6 consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58-minute long Earth
occultations every 94 minutes.

The observations applied space craft rolls each two orbits to allow
roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbit 1, 3, and 5, and those for 2, 4, and 6
differed by $25^{\circ}$. At the separation of 2M1207b, this angle
difference corresponds to a displacement of $0.34''$, or 2.75 and 2.30
resolution elements in F125W and F160W, respectively. In each orbit we
took 8 SPARS10 exposure sequence with NSAMP=10, alternating between
F160W and F125W filters, with 2--3 identical exposures in each
exposure sequence. To improve sampling and reduce the risk that the
core of point spread function (PSF) is affected by bad pixels, we
applied a 4-point dither pattern with differential ``X/Y'' offsets of
1.375" in the detector frame, providing optimal non-integral (half
pixel) step of 10.5 and 8.5 pixels in F125W and F160W,
respectively. In total, we obtained 70 and 64 images with 10 non-destructive
read-outs in F125W and F160W, respectively.

@@

We started the reduction from the \flt{} produced by the WFC3's
\texttt{calwfc3} pipeline. We did not opt to use \ima{} that contain
all non-destructive read-outs, because they provided less information
on 2M1207A, which saturated after the first few samples.  The \flt{}
are results of basic calibration, including dark current correction,
non-linearity correction, flat field correction, as well as
up-the-ramp fit on the non-destructive read-outs. 
Pixels with data quality flags ``bad detector pixels'', ``unstable
response'', and ``bad or uncertain flat value'' were masked out and
excluded from further analysis as suggested by previous transit
exoplanet spectroscopic observations \citep[e.g.][]{Berta2012,
  Kreidberg2014}.

The major challenge of high contrast observation with WFC3/IR is the
fact that the detector is significantly under-sampled.  2M1207A and b
are only separated by $\sim6$ pixels or $\sim$5 FWHM of the PSF on the
detector. When applying roll subtraction, notable artifacts are
introduced by image shifting and interpolation.  \tinytim{} PSF
simulator \citep{Krist1995} offers a solution by providing Nyquist or
better sampled PSF, but systematic errors of \tinytim{} PSF for WFC3
limits its ability in high precision photometry \citep{Biretta2014}.
Building on the large number of PSFs obtained in our program at two
different roll angles, we followed a novel, two-step approach that
uses a hybrid PSF.  First, based on our observations, we derived
correction maps for \tinytim{} that accurately described the scattered
light component for the primary at the correct location on the
detector. Second, we carried out a PSF-photometry using hybrid PSFs
composed by \tinytim{} PSFs and the correction map by simultaneously
minimizing the residuals from the primary and the secondary.
% However, we are able to fully characterize the difference of model
% and observed PSFs with 6 orbits time-resolved observation data. To
% obtain robust \tinytim{} PSF photometry, we design a 2-round PSF
% fitting strategy: 1. calculating correction map for \tinytim{};
% 2. hybrid PSF photometry.

For both of 2 steps, we used \tinytim{} to calculate 10$\times$
over-sampled model PSFs based on the filters, the spectra (2M1207A:
\cite{Bonnefoy2014}, 2M1207b: \cite{Patience2010}), the telescope's
actual focus, and the telescope jitter.  We used the set of \tinytim{}
parameters provided by \cite{Biretta2014} to improve modeling the cold
mask, diffraction spikes, and the coma. The focus parameters are 
interpolated to the precise time of the observations using the
tabulated values provided by
STScI\footnote{\url{http://www.stsci.edu/hst/observatory/focus/FocusModel}}.
To align the \tinytim{} PSF to the observed PSF of 2M1207A, we
moved the over-sampled PSF on a coordinate grid (gird size=0.001 pixel)
using cubic interpolation, and searched for the position that minimizes
the rms difference of the observed and the re-binned \tinytim{} PSF over
a region centered on 2M1207A with a 5-pixel-radius aperture centered
on 2M1207b excluded.  Then we introduced another \tinytim{} PSF for
2M1207b and fit the position of 2M1207b and the scales of the
\tinytim{} PSFs of 2M1207A and b simultaneously by minimizing the
residual from both primary and secondary. In the first step, we discovered that the
difference of observed PSFs and model PSFs was very stable for a
specified telescope roll angle and dithering position. Therefore, at the end of the first step,
we derive 8 (2 roll angles $\times$ 4 dithering positions) correction
maps for each filter:
\begin{equation}
  \mathrm{Corr = Median(PSF_{obs.} - PSF_{model} )}
\end{equation}
where $\mathrm{PSF_{model}}$ was a combination of two scaled \tinytim{} PSFs
for 2M1207A and b. In the second round, we combined the correction
map linearly with the two \tinytim{} PSFs to generate hybrid PSFs,
and scaled the correction map together with the two PSFs so that the
residual, which is expressed as
\begin{equation}
  \mathrm{residual} = \mathrm{Image} - a\cdot \mathrm{PSF_{A}} - b\cdot
    \mathrm{PSF_{b}}- c\cdot\mathrm{ Corr}
  \end{equation}
  is minimized by least square fitting.
We found that by introducing the correction term,
the reduced $\chi^{2}$ was decreased from $\sim 10$ to
$\sim 1$. Relative photometry was acquired from the scaling
parameters of the \tinytim{} PSFs.

Our final step was to correct for a slight apparent trend between the
position of the targets on the detector and their flux. We attributed
this to a combination of slight changes in the PSF profile due to
pixelation and to the effect of imperfectly corrected pixel-to-pixel
sensitivity variations. We corrected for these apparent
position-dependent flux changes by normalizing each photometric point
by the median of all fluxes measured when the target was at the
same position, i.e. combining data over 6 orbits.  We note that this
correction was small and, as we demonstrate in the next sections,
could not introduce artificial modulations that resemble the long-period
variations that we identified in 2M1207b.

% PSF profiles change with exposure positions due to
% pixelation, especially for the case that WFC3 IR is significantly
% under-sampled. Also, the flat fields may potentially have large scale
% structures \citep{dressel2012wide}. We find a correlation of
% photometry with PSF positions on detector frame for both 2M1207A and
% b. Correction is made by normalizing each group of
% exposures that have the same dithering position and telescope roll angle
% individually -- we take the median of the fluxes that are measured
% from these exposures as normalization factors and divided them from
% every photometric measurement. Because the normalization factor for
% each group of exposures is calculated across the whole observation in
% time domain,
% this step has negligible impacts on variability
% analysis.

@@

First we estimated the photon noise for the photometry of
2M1207b. The total photon noise of the photometry was calculated by
combining the photon noise of every pixel, which was derived
from count rates and detector gain. The photon noises 
in F125W and F160W are 1.33\% and 1.02\%, respectively.

Since the PSFs for the 2M1207A and b were fitted simultaneously, the
uncertainties for photometry and position of the primary and secondary
were coupled. Errors in position measurements of 2M1207A could
potentially affect the photometry of 2M1207b. We used a Monte Carlo
(MC) method to evaluate the overall systematics of the PSF fitting. We
applied photometry to images that were added with random
Poisson noise and repeated the photometry procedure for 1000 times. The
uncertainties for F125W and F160W photometry were found to be 1.34\% and 1.12\%,
respectively.

@@

A further contribution to photometric uncertainties may be introduced
by imperfectly corrected pixel-to-pixel sensitivity
differences. 2M1207b were observed at 8 different positions on the
detector (2 rolls $\times$ 4 dithering positions). Imperfect flat
field correction could introduce position-dependent
differences in the count rates. The uncertainty of WFC3 IR flat field
is typically $\sim 1\%$ \citep{dressel2012wide}.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
    variability verification tests. Individual measurements are
    plotted with gray crosses. Photometric measurements of the same exposure
    sequence are binned, and binned photometry are plotted with points
    or squares. Best fitted sinusoids are plotted with solid
    lines. {\em Upper}: binned measurements taken in dithering
    position 1 and 3 (red points) and that taken in 2 and 4 (blue
    squares) are plotted with different symbols. They demonstrate same
    trend of modulation. In upper left panel, the green line a sinusoid
    fitted with all parameters freely varying, and the purple line is
    a sinusoid fitted with the period set the same as that of
    F125W. {\em Middle}: sinusoids fitted without using the
    data taken in Orbit \#1. These curves are almost identical to the
    curves plotted in upper panel. {\em Lower}: photometry measured
    with AFEM-added images and best fitted sinul curves. These
    points and curves are also almost identical to those plotted
    in the upper panel.}
  \label{fig:2}
\end{figure*}

@@

The light curves that resulted from our photometry showed apparently
sinusoidal modulations, discussed in more detail in
\S\ref{Results}. To verify that these modulations are intrinsic to
the object and not the result of our data reduction procedures or 
instrumental changes, we carried out three different tests.

@@

As a second test, we repeated the analysis neglecting the first
orbit. The motivation behind this test is that, due to spacecraft
thermal settling, the first orbits of HST observations are often
slightly unstable, and are neglected in high-precision studies
\citep[e.g.][]{Mandell2013}. Indeed, in our analysis, 2M1207A is
significantly fainter in the first orbit (Figure \ref{fig:3}) than in
the subsequent ones.  Our analysis based on orbits 2--6 found
essentially identical results to our analysis using the whole 6 orbits, based on
which we conclude that the first less reliable orbit does not affect
our results significantly (Figure \ref{fig:2}, middle panel).

As a third test, we explored whether a subset of images, perhaps due to
imperfect normalization or correlation with specific instrument states,
could drive the light curves into apparently sinusoidal shapes. To
test this possibility, we split the data into two temporally
overlapping halves: subset one were images taken at dithering position
1 and 3; subset two were those taken at dithering position 2 and
4. For both subsets, we repeated our analysis independently.  For both
of F125W and F160W, two halves demonstrated similar sinusoidal
modulations.  Our analysis detect sinusoidal modulations in {\em both}
subsets and in {\em both} filters, with periods and amplitudes
consistent with those derived from the complete data set (Figure
\ref{fig:2}, upper panel).

@@

We used a MC method to analyze the light curve, and provided amplitudes
and periods as well as their uncertainties for both filters. We
generated series of Gaussian distributed random noise with standard deviation
same as the photon noise, added the noise to the original light curves, and
fitted sinusoids to the newly combined light curves. We repeated these
procedures  for 100,000 times and obtain the distributions of period
and amplitude (Figure \ref{fig:4}).
